---
title: "Week5-Intro to Probability and Data: "
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---
## Exploring the BRFSS data
#### by Soyoung Park
#### May 26, 2017
## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data

(3 points) Describe how the observations in the sample are collected, and the implications of this data collection method on the scope of inference (generalizability / causality). Note that you will need to look into documentation on the BRFSS to answer this question. See http://www.cdc.gov/brfss/ as well as the "More information on the data" section below.

* 2 pt for correct reasoning for generabizability – Answer should discuss whether random sampling was used. Learners might discuss any reservations, those should be well justified.

> **ANSWER**:  
BRFSS completes more than 400,000 adult interviews each year, making it the largest continuously conducted multi-mode (mail, landline phone, and cell phone) health survey system in the world. Home telephone numbers are obtained through random-digit dialing [Ref](https://www.cdc.gov/brfss/factsheets/pdf/brfss-history.pdf). 


* 1 pt for correct reasoning for causality – Answer should discuss whether random assignment was used.

> **ANSWER**:  
However, care must be taken to draw any conclusion with statistical significance. For example, according to the report regarding [limitation of health care access(HCA) module](https://www.cdc.gov/brfss/data_documentation/pdf/2013-2014_hcs.pdf), "data are self-reported, which could result in recall and social desirability response biases".

* * *

## Part 2: Research questions

 (11 points) Come up with at least three research questions that you want to answer using these data. You should phrase your research questions in a way that matches up with the scope of inference your dataset allows for. Make sure that at least two of these questions involve at least three variables. You are welcomed to create new variables based on existing ones. With each question include a brief discussion (1-2 sentences) as to why this question is of interest to you and/or your audience.
 

* 3 pts for each research question (x 3)
    * Should be phrased in a non-causal way (1 pt)
    * Should be well defined / not vague (1 pt)
    * Is clear why this is of interest to the author / audience (1 pt)
* 2 pts for at least two of the questions involving three variables

**Research quesion 1:**  

> **ANSWER**:   
Is there any correlation between depression and the following factors? Rank out p-value to evaluate how strongly correlated.

> * *menthlth*: Number Of Days Mental Health Not Good  
  Now thinking about your mental health, which includes stress, depression, and problems with emotions, for how many days during the past 30 days was your mental health not good?

> * *poorhlth*: Poor Physical Or Mental Health  
  During the past 30 days, for about how many days did poor physical or mental health keep you from doing your usual activities, such as self-care, work, or recreation?

> * *sleptim1*: How Much Time Do You Sleep  
  I would like to ask you about your sleep pattern. On average, how many hours of sleep do you get in a 24-hour period?

> * *fruit1*: How Many Times Did You Eat Fruit?
  *During the past month, not counting juice, how many times per day, week, or month did you eat fruit? Count fresh, frozen, or canned fruit.(Read only if necessary: “Your best guess is fine. Include apples, bananas, applesauce, oranges, grape fruit, fruit salad, watermelon, cantaloupe or musk melon, papaya, lychees, star fruit,)*

> * *fvbeans*: How Many Times Did You Eat Beans Or Lentils?
  During the past month, how many times per day, week, or month did you eat cooked or canned beans, such as refried, baked, black, garbanzo beans, beans in soup, soybeans, edamame, tofu or lentils. Do NOT include long green beans.(Read only if necessary: “Include round or oval beans or peas such as navy, pinto, split peas, cow peas, garbanzo beans, lentils, soy beans and tofu. Do NOT include long green beans such as string beans, broad or winged beans, or pole beans.”)

> * *fvgreen*: How Many Times Did You Eat Dark Green Vegetables?
During the past month, how many times per day, week, or month did you eat dark green vegetables for example broccoli or dark leafy greens including romaine, chard, collard greens or spinach?

> * *fvorang*: How Many Times Did You Eat Orange-Colored Vegetables?
  During the past month, how many times per day, week, or month did you eat orange-colored vegetables such as sweet potatoes, pumpkin, winter squash, or carrots? (Read only if needed: “Winter squash have hard, thick skins and deep yellow to orange flesh. They include acorn, buttercup, and spaghetti squash.”)

> * *vegetab1*: How Many Times Did You Eat Other Vegetables?
  Not counting what you just told me about, during the past month, about how many times per day, week, or month did you eat OTHER vegetables? Examples of other vegetables include tomatoes, tomato juice or V-8 juice, corn, eggplant, peas, lettuce, cabbage, and white potatoes that are not fried such as baked or mashed potatoes.(Read only if needed: “Do not count vegetables you have already counted and do not include fried potatoes.”)

> * *exerany2*: Exercise In Past 30 Days
  During the past month, other than your regular job, did you participate in any physical activities or exercises such as running, calisthenics, golf, gardening, or walking for exercise?

> * *exract11*: Type Of Physical Activity
  What type of physical activity or exercise did you spend the most time doing during the past month?

> * *exeroft1*: How Many Times Walking, Running, Jogging, Or Swimming
  How many times per week or per month did you take part in this activity during the past month?

> * *exerhmm1*: Minutes Or Hours Walking, Running, Jogging, Or Swimming
  And when you took part in this activity, for how many minutes or hours did you usually keep at it?

> * *exract21*: Other Type Of Physical Activity Giving Most Exercise During Past Month
  What other type of physical activity gave you the next most exercise during the past month?

> * *exeroft2*: How Many Times Walking, Running, Jogging, Or Swimming
  How many times per week or per month did you take part in this activity during the past month?

> * *strength*: How Many Times Did You Do Physical Activities Or Exercises To Strengthen
  During the past month, how many times per week or per month did you do physical activities or exercises to STRENGTHEN your muscles? [Do NOT count aerobic activities like walking, running, or bicycling. Count activities using your own body weight like yoga, sit-ups or push-ups and those using weight machines, free weights, or elastic bands.]


**Research quesion 2:**  

> **ANSWER**:  
  How do people with joint problem likely to manage? What do people do and what doctor suggest?

> * *joinpain*: How Bad Was Joint Pain
  Please think about the past 30 days, keeping in mind all of your joint pain or aching and whether or not you have taken medication. DURING THE PAST 30 DAYS, how bad was your joint pain ON AVERAGE? [Please answer on a scale of 0 to 10 where 0 is no pain or aching and 10 is pain or aching as bad as it can be.]

> * *arthsocl*: Social Activities Limited Because Of Joint Symptoms
  During the past 30 days, to what extent has your arthritis or joint symptoms interfered with your normal social activities, such as going shopping, to the movies, or to religious or social gatherings?

> * *lmtjoin3*: Limited Because Of Joint Symptoms
  Arthritis can cause symptoms like pain, aching, or stiffness in or around a joint. Are you now limited in any way in any of your usual activities because of arthritis or joint symptoms?

> * *arthdis2*: Does Arthritis Affect Whether You Work
  Are you now limited in any way in any of your usual activities because of arthritis or joint symptoms? Do arthritis or joint symptoms now affect whether you work, the type of work you do or the amount of work you do?

> * *arttoday*: What Can You Do Because Of Arthritis Or Joint Symptoms
  Earlier you indicated that you had arthritis or joint symptoms.Thinking about your arthritis or joint symptoms, which of the following best describes you TODAY?

> * *arthwgt*: Dr. Suggest Lose Weight For Arthritis Or Joint Symptoms
  Has a doctor or other health professional EVER suggested losing weight to help your arthritis or joint symptoms?

> * *arthexer*: Dr. Suggest Use Of Physical Activity Or Exercise For Arthritis Or Joint Symptoms
  Has a doctor or other health professional EVER suggested physical activity or exercise to help your arthritis or joint symptoms? (If the respondent is unclear about whether this means an increase or decrease in physical activity, this means increase.)

> * *arthedu*: Ever Taken Class In Managing Arthritis Or Joint Symptoms
  Have you EVER taken an educational course or class to teach you how to manage problems related to your arthritis or joint symptoms?



**Research quesion 3:**
> **ANSWER**:
Do people with joint problem tent to have depression?


## Part 3: Exploratory data analysis

(30 points) Perform exploratory data analysis (EDA) that addresses each of the three research questions you outlined above. Your EDA should contain numerical summaries and visualizations. Each R output and plot should be accompanied by a brief interpretation.

10 pts for each for each research question (x 3)

* 3 pts for plots
    * Plots should address the research questions (1 pt)
    * Plots should be constructed correctly (1 pt)
    * Plots should be formatted well – size not too large, not too small, etc. (1 pt)
* 3 pts for summary statistics
    * Summary statistics should address the research questions (1 pt)
    * Summary statistics should be calculated correctly (1 pt)
    * Summary statistics should be formatted well – not taking up pages and pages, etc. (1 pt)
* 4 pts for narrative
    * Each plot and/or R output should be accompanied by a narrative (1 pt)
    * Narrative should interpret the visuals / R output correctly (2 pts)
    * Narrative should address the research question (1 pts)

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.


for multivariate plot, 


```{r load_libraries, message=FALSE}
#install.packages("car")
library("car")
```

**Research quesion 1:**

```{r}

```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```



### Review criterialess 

* Overall (6 points)
    * Uploaded HTML document resulting from the Rmd template: 1 pt
    * Organization: 1 pts
    * Readability of the text: 2 pts
    * Readability of the code: 2 pts

* Do you understand the research question?
* Is it clear how data can be used to answer this research question as its phrased?
* Scope of inference - generalizability: Is the data collection explained clearly? Did the writer correctly identify the population of interest? Did the writer correctly decide whether the findings from this analysis can be generalized to that population, or, if not, a subsection of that population? Is their explanation satisfactory to make this decision? Are potential sources of bias discussed, and if so, is the discussion satisfactory?
* Scope of inference - causality: Did the writer identify correctly whether these data can be used to establish causal links between the variables of interest. Is the explanation satisfactory?
* Are the research questions well defined and is it clear why they are of interest to the author / audience? Are appropriate numbers of variables used in each research question?
* Are appropriate summary statistics calculated, and are they explained/interpreted in context of the data and the research question?
* Are appropriate visualizations included, and are they explained/interpreted in context of the data and the research question?
* Did the writer address what the findings from the exploratory analysis suggest about the research question?
